SPP-Perc: Peceptron Learning Based Enhanced Signature Path Prefetcher

### 2019-04-17 abs.tex
Paragraph spacing

  In this paper, we introduce Perceptron-based Prefetch Filtering
  (PPF) as a way to increase the coverage of the prefetches generated by
  a baseline prefetcher without negatively impacting accuracy. PPF
  enables more aggressive tuning of a given baseline prefetcher,
  leading to increased coverage by filtering out the growing numbers
  of inaccurate prefetches such an aggressive tuning implies. We also
  explore a range of features to use to train PPF's perceptron layer
  to identify inaccurate prefetches. PPF improves performance on a
  memory-intensive subset of the SPEC CPU 2017 benchmarks by 3.78\%
  for a single-core configuration, and by 11.4\% for a 4-core
  configuration, compared to the baseline prefetcher alone.

>>>>

  In this paper, we introduce Perceptron-based Prefetch Filtering
  (PPF) as a way to increase the coverage of the prefetches generated by
  an underlying prefetcher without negatively impacting accuracy. PPF
  enables more aggressive tuning of the underlying prefetcher,
  leading to increased coverage by filtering out the growing numbers
  of inaccurate prefetches such an aggressive tuning implies. We also
  explore a range of features to use to train PPF's perceptron layer
  to identify inaccurate prefetches. PPF improves performance on a
  memory-intensive subset of the SPEC CPU 2017 benchmarks by 3.78\%
  for a single-core configuration, and by 11.4\% for a 4-core
  configuration, compared to the underlying prefetcher alone.

### 2019-04-17 back.tex
Paragraph spacing

\subsection{{\color{red}Underlying Prefetcher: SPP}}
\label{Background-SPP}

Kim {\em et. al.} proposed Signature Path Prefetcher (SPP)~\cite{SPP},
a confidence-based lookahead prefetcher.  SPP creates a signature
associated with a page address by compressing the history of
accesses. By correlating the signature with future likely delta
patterns, SPP learns both simple and complicated memory access
patterns quickly.  While the basic idea of perceptron based prefetch
filtering is applicable to any lookahead prefetcher, we develop a
practical implementation of our proposed perceptron prefetch filter
using SPP as our {\color{red}underlying mechanism}. 
Here we describe the basic architecture of SPP.

\noindent \textbf{Signature Table:} As shown on the left side of
Figure~\ref{fig:spp_update}, the Signature Table keeps track of 256
most recently accessed pages. It is meant to capture memory access
patterns within a page boundary. SPP indexes into an entry of the
Signature Table using the page number. For each entry corresponding to
a page, the table stores a `last block offset' and an `old
signature'. Last block offset is the block offset of the last memory
access of that given page. The block offset is calculated with respect
to the page boundary. The signature is a 12-bit compressed
representation of the past few memory accesses for that page. The
signature is calculated as:
$$New Signature = (\,Old Signature << 3 bits\,) \;\;XOR\;\;
(\,Delta\,)$$ Delta is the numerical difference between the block
offset of the current and the previous memory access. In case a
matching page entry is found, the stored signature retrieved and used
to index into the Pattern Table.  This process is illustrated in
Figure~\ref{fig:spp_update}.

\noindent \textbf{Pattern Table:} The Pattern Table, shown on the right
side in Figure~\ref{fig:spp_update} is indexed by the signature
generated from the Signature Table.  Pattern Table holds predicted
delta patterns and their confidence estimates. Each entry indexed by
the signature holds up to 4 unique delta predictions.

% djimenez: it does look like we're pushing it with all the stuff about SPP.
% Let's describe it enough to give the reader the idea that he/she has a vague
% understanding of how it works and can refer to the MICRO paper for more
% details. (I'm not saying cut anything that's already there, just don't expand.)

%% NOTE: Can we do away with this example? 
%% If it looks like we are writing too much about SPP
%% EDIT: Decided to reorganize and remove this for now.
%% 	Consider the case that incoming page number 10 with a block offset 3
%% 	finds a match in the ST. The retrieved pattern signature is 0x30 and
%% 	the Last Offset is 1. Since now there is the Last Block Offset (1),
%% 	Incoming Block Offset (3), Old Signature (0x30), and New Signature
%% 	calculated as per above equation (0x182), SPP can infer
%% 	non-speculatively that the given pattern of memory accesses (as
%% 	captured in the Old Signature) leads to the particular Delta. In
%% 	general, Delta is defined as the difference between the prefetch
%% 	suggestion block and the initial block which triggered the prefetch.
%% 	In this case, since SPP is in learning phase, it is defined as the
%% 	difference between the Incoming Block Offset and the Last Block Offset
%% 	(+2 in this case). This, in turn generates the new memory pattern
%% 	(New Signature). This newly learned signature pattern and the delta
%% 	is stored in the Pattern Table.

%  -- This looks like a detail that is confusing and not necessary to
%  understand SPP:
% \noindent \textbf{Prefetch Table:} The Prefetch Table is a 1024 entry
% direct-mapped table that keeps a record of last few entries
% prefetched.  This proves useful for updating the prefetcher state if a
% tracked prefetch leads to a demand hit or a cache
% eviction. \textit{Note:} The original paper refers to this as the
% Prefetch Filter but since in our context filter refers to the
% perceptron filter, we will be calling this structure as the Prefetch
% Table to avoid any confusion.

\noindent \textbf{Lookahead Prefetching:} On each trigger, SPP goes
down the program speculation path using its own prefetch suggestion.
Using the current prefetch as a starting point, it re-accesses the Pattern
Table to generate further prefetches.  As illustrated in
Figure~\ref{fig:spp_structure}, it repeats the cycle of accessing
Pattern Table and updating the signature based on highest confidence
prefetch from the last iteration.  The iteration counter on which SPP
manages to predict prefetch entries in the lookahead manner is
characterized as its `depth'. While doing so, SPP also keeps
compounding the confidence in each depth. Thus as depth increases,
overall confidence keeps decreasing.

\noindent \textbf{Confidence Tracking}: As shown in
Figure~\ref{fig:spp_structure}, the Pattern Table keeps a count of
hits to each signature through a counter C\textsubscript{sig}. The
number of hits for a given delta per signature are tracked using a
counter C\textsubscript{delta}.  The confidence for a given delta is
approximated through C\textsubscript{d} = C\textsubscript{delta} /
C\textsubscript{sig}.  When SPP enters into a lookahead mode, the path
confidence P\textsubscript{d} is given as:
$$P\textsubscript{d} \;=\; \alpha  \;.\;  C\textsubscript{d}  \;.\;
P\textsubscript{d-1}$$ Here $\alpha$ represents the global accuracy,
calculated as the ratio of the number of prefetches which led to a
demand hit to the number of prefetches recommended in total. The range
of $\alpha$ is [0,1].  The lookahead depth is represented by $d. For $d = 1$, when
SPP is in non-speculative mode, P\textsubscript{0} can be thought of
as 1.  The final P\textsubscript{d} is thresholded against prefetch
threshold (T\textsubscript{p}) to reject the low confidence
suggestions and then against a numerically bigger fill threshold
(T\textsubscript{f}) to decide whether to send the prefetch to
L2 Cache (high confidence prefetch) or Last Level Cache (low
confidence prefetch). The two thresholds were empirically set to 25
and 90 respectively, on the scale of 0 to 100.

>>>>

\subsection{Underlying Prefetcher: SPP}
\label{Background-SPP}

Kim {\em et. al.} proposed Signature Path Prefetcher (SPP)~\cite{SPP},
a confidence-based lookahead prefetcher.  SPP creates a signature
associated with a page address by compressing the history of
accesses. By correlating the signature with future likely delta
patterns, SPP learns both simple and complicated memory access
patterns quickly.  While the basic idea of perceptron based prefetch
filtering is applicable to any lookahead prefetcher, we develop a
practical implementation of our proposed prefetch filter
using SPP as our underlying mechanism. 
Here we describe the basic architecture of SPP.
\newline
\newline
\noindent \textbf{Signature Table:} As shown on the left side of
Figure~\ref{fig:spp_update}, the Signature Table (ST) keeps track of 256
most recently accessed pages. It is meant to capture memory access
patterns within a page boundary. SPP indexes into an entry of the
Signature Table using the page number. For each entry corresponding to
a page, the table stores a `last block offset' and an `old
signature'. Last block offset is the block offset of the last memory
access of that given page. The block offset is calculated with respect
to the page boundary. The signature is a 12-bit compressed
representation of the past few memory accesses for that page. The
signature is calculated as:
$$New Signature = (\,Old Signature << 3 bits\,) \;\;XOR\;\;
(\,Delta\,)$$ Delta is the numerical difference between the block
offset of the current and the previous memory access. In case a
matching page entry is found, the stored signature retrieved and used
to index into the Pattern Table.  This process is illustrated in
Figure~\ref{fig:spp_update}.
\newline
\newline
\noindent \textbf{Pattern Table:} The Pattern Table (PT), shown on the right
side in Figure~\ref{fig:spp_update} is indexed by the signature
generated from the Signature Table.  Pattern Table holds predicted
delta patterns and their confidence estimates. Each entry indexed by
the signature holds up to 4 unique delta predictions.
%
% djimenez: it does look like we're pushing it with all the stuff about SPP.
% Let's describe it enough to give the reader the idea that he/she has a vague
% understanding of how it works and can refer to the MICRO paper for more
% details. (I'm not saying cut anything that's already there, just don't expand.)
%
%% NOTE: Can we do away with this example? 
%% If it looks like we are writing too much about SPP
%% EDIT: Decided to reorganize and remove this for now.
%% 	Consider the case that incoming page number 10 with a block offset 3
%% 	finds a match in the ST. The retrieved pattern signature is 0x30 and
%% 	the Last Offset is 1. Since now there is the Last Block Offset (1),
%% 	Incoming Block Offset (3), Old Signature (0x30), and New Signature
%% 	calculated as per above equation (0x182), SPP can infer
%% 	non-speculatively that the given pattern of memory accesses (as
%% 	captured in the Old Signature) leads to the particular Delta. In
%% 	general, Delta is defined as the difference between the prefetch
%% 	suggestion block and the initial block which triggered the prefetch.
%% 	In this case, since SPP is in learning phase, it is defined as the
%% 	difference between the Incoming Block Offset and the Last Block Offset
%% 	(+2 in this case). This, in turn generates the new memory pattern
%% 	(New Signature). This newly learned signature pattern and the delta
%% 	is stored in the Pattern Table.
%
%  -- This looks like a detail that is confusing and not necessary to
%  understand SPP:
% \noindent \textbf{Prefetch Table:} The Prefetch Table is a 1024 entry
% direct-mapped table that keeps a record of last few entries
% prefetched.  This proves useful for updating the prefetcher state if a
% tracked prefetch leads to a demand hit or a cache
% eviction. \textit{Note:} The original paper refers to this as the
% Prefetch Filter but since in our context filter refers to the
% perceptron filter, we will be calling this structure as the Prefetch
% Table to avoid any confusion.
\newline
\newline
\noindent \textbf{Lookahead Prefetching:} On each trigger, SPP goes
down the program speculation path using its own prefetch suggestion.
Using the current prefetch as a starting point, it re-accesses the Pattern
Table to generate further prefetches.  As illustrated in
Figure~\ref{fig:spp_structure}, it repeats the cycle of accessing
the PT and updating the signature based on highest confidence
prefetch from the last iteration.  The iteration counter on which SPP
manages to predict prefetch entries in the lookahead manner is
characterized as its `depth'. While doing so, SPP also keeps
compounding the confidence in each depth. Thus as depth increases,
overall confidence keeps decreasing.
\newline
\newline
\noindent \textbf{Confidence Tracking}: As shown in
Figure~\ref{fig:spp_structure}, the Pattern Table keeps a count of
hits to each signature through a counter C\textsubscript{sig}. The
number of hits for a given delta per signature are tracked using a
counter C\textsubscript{delta}.  The confidence for a given delta is
approximated through C\textsubscript{d} = C\textsubscript{delta} /
C\textsubscript{sig}.  When SPP enters into a lookahead mode, the path
confidence P\textsubscript{d} is:
$$P\textsubscript{d} \;=\; \alpha  \;.\;  C\textsubscript{d}  \;.\;
P\textsubscript{d-1}$$ Here $\alpha$ represents the global accuracy,
calculated as the ratio of the number of prefetches which led to a
demand hit to the number of prefetches recommended in total. The range
of $\alpha$ is [0,1].  The lookahead depth is represented by $d$. 
For $d = 1$, when
SPP is in non-speculative mode, P\textsubscript{0} can be thought of
as 1.  The final P\textsubscript{d} is thresholded against prefetch
threshold (T\textsubscript{p}) to reject the low confidence
suggestions and then against a numerically bigger fill threshold
(T\textsubscript{f}) to decide whether to send the prefetch to
L2 Cache (high confidence prefetch) or Last Level Cache (low
confidence prefetch). The two thresholds were empirically set to 25
and 90 respectively, on the scale of 0 to 100.

### 2019-04-17 conc.tex
Paragraph spacing

\section{Conclusion}
\label{Conclusion}
In this paper, we introduce the Perceptron-Based Prefetch Filtering
(PPF).  PPF acts as an independent check on the quality of predictions
made by the underlying prefetch engine.  We also created a case study
implementation of PPF using SPP as the baseline prefetch engine, while
in principle other prefetchers could be used.  We show that PPF
effectively filters bad prefetches, such that the given underlying
prefetcher can be highly aggressively tuned to achieve increasing
coverage.  PPF improves performance over the base prefetcher by up to
11.4\%.  PPF is a robust and adaptable technique that can be used to
enhance any existing prefetcher and can be a valuable tool in the
design of future memory latency constrained systems.

>>>>

\section{Conclusion}
\label{Conclusion}
In this paper, we introduce the Perceptron-Based Prefetch Filtering
(PPF).  PPF acts as an independent check on the quality of predictions
made by the underlying prefetch engine.  We also created a case study
implementation of PPF using SPP as the underlying prefetcher, while
in principle other prefetchers could be used.  We show that PPF
effectively filters bad prefetches, such that the given underlying
prefetcher can be highly aggressively tuned to achieve increasing
coverage.  PPF improves performance over the underlying 
prefetcher by up to 11.4\%.  PPF is a robust and adaptable technique 
that can be used to enhance any existing prefetcher and can be a 
valuable tool in the design of future memory latency constrained systems.

### 2019-04-17 design_arch.tex
merging
Paragraph spacing

Perceptron learning is a light-weight mechanism to pull together
disparate forms of information and synthesize a decision from
them. Our work considers a number of features corresponding to a
prefetch, such as speculation depth, page address and offset, and uses
this information as the inputs to our perceptron-based filter in order
to predict the usefulness of a prefetch.  Here, we discuss our the
design of our proposed perceptron prefetch filter (PPF).  PPF enhances
{\color{red}an underlying prefetcher} by filtering out predicted unused prefetches. 
PPF is a generalized prefetch filtering mechanism that may be adapted 
to any prefetcher with appropriate feature selection and modifications 
which we describe below.

@@

Figure~\ref{fig:PPF_Datapath} shows the microarchitecture of PPF, as
well as the steps required to filter out not-useful prefetches. The
perceptron filter is organized as a set of tables, where each entry in
the tables holds a \textit{weight}.  For a configuration of PPF using
$N$ number of features, $N$ different tables of weights are needed.
Each feature is used to index a distinct table. The number of entries
of each table varies according to the corresponding feature, hence,
different number of bits are needed to index different tables. Each
weight is a 5-bit saturating counter ranging from -16 to +15. We found
that having 5-bit weights provides a good trade-off between accuracy
and area. A detailed explanation of the storage overhead of PPF can
be found in Section~\ref{Method-Overheads}.

%%\pg{the text below would be easier to follow if you used the figure
%%  more. make reference to the figure at the points that you are
%%  talking about}

\noindent \textbf{Inferencing}\\
%ETG:is this like an explanation of how prefetching happens? 
%EB: I see this as a short context of the datapath leading to the perc filter.
The {\color{red}prefetcher} is triggered on every demand access to the L2
Cache, as seen in Figure~\ref{fig:PPF_Hierarchy}.  At this point, 
it has the opportunity to trigger a prefetch.
If it does so, it will also need to decide how many cache blocks to
prefetch.  These blocks can be either placed in the L2 or L3 cache
according to the confidence of the prefetching mechanism.  Once the
{\color{red}underlying} prefetcher is triggered, the suggested prefetch candidates are
fed to the perceptron filter to determine the usefulness of these
prefetches. The filter ultimately decides whether to issue the
prefetch suggestions of the {\color{red}underlying} prefetcher. As shown in step 1
of Figure~\ref{fig:PPF_Datapath}(a), to make the decision,
each feature corresponding to a suggested prefetch is used to index a
table and all the corresponding weights are summed. The sum denotes
the confidence value for the suggested prefetch, and is thresholded
against two different values: $\tau_{hi}$ and $\tau_{lo}$.

Prefetches whose sum exceeds $\tau_{hi}$ are placed into the L2 cache.
The higher confidence value hints the prefetch would be useful and
should be prioritized.  A prefetch for which the features result in a
confidence value between $\tau_{lo}$ and $\tau_{hi}$ is allocated in
the larger LLC, as the filter is moderately confident of the future
reuse of the cache block, but not enough to possible pollute a
significantly smaller L2.  Suggested prefetches for which the features
lead to a confidence value lower than $\tau_{lo}$ are not prefetched,
as the low confidence value represents that the perceptron learned
that a similar set of features are associated with non-useful
prefetches.

\noindent \textbf{Recording}\\
As shown in step 2 of Figure~\ref{fig:PPF_Datapath}(a) the prefetches
that make it through the inference stage are recorded in the
``Prefetch Table''. The prefetch table is a 1,024-entry, direct mapped
structure that contains all metadata required to re-index the
perceptron entries for training. {\color{red}Ten bits of the address are 
used to index into the tables, and another six bits are stored to 
perform tag matching.}

In addition to the prefetch table mentioned above, PPF also maintains
a 1,024-entry direct-mapped ``Reject Table.'' If a prefetch suggestion
is rejected by the perceptron layer, it is logged into the reject
table. The table is used to train the perceptron to avoid false
negatives \textit{i.e.}, cases where the prediction suggested to
reject the prefetch but the prefetch was proven to be useful based on
the observed demand accesses to the L2.

\noindent \textbf{Feedback and Data Retrieval}\\
As depicted in step 3, when there is an eviction or a demand access 
to the L2, training for both the {\color{red}underlying} prefetcher 
and our filter mechanism is triggered. The address of the cache block 
that triggered the training is used to index both the Prefetch and 
Reject tables. {\color{red}If it is a match, the corresponding
features are retrieved to index into the tables of perceptron weights.}

% Which bits? are the tables accesses sequentially?

\noindent \textbf{Training}\newline As can be seen in step 4 of, 
the address from the demand
request triggering the training is looked up in both tables. If the
address is in the prefetch table and marked as valid, this hints the
previous prediction was correct and this is a useful prefetch. We
compute the sum of the corresponding weights. If the sum falls below a
specific threshold, training occurs and the corresponding weights are
adjusted accordingly. These thresholds are introduced in order to
avoid over-training, helping the filter adapt quickly to changes in
memory behavior. These thresholds are referred to as $\theta_p$ and
$\theta_n$, respectively for the positive and negative values of
training saturation.

@@

Parallel to accessing the prefetch table, on a demand access, the
reject table is accessed. Before the demand access triggers the next
set of prefetches, the reject table is checked for a valid entry. A
hit means that the corresponding cache block was initially suggested
by the {\color{red}underlying} prefetcher, but wrongly rejected by the perceptron
filter. The perceptron filter learns from this and makes use of the
corresponding features associated to the original prefetch request,
which are stored in the reject table, to index the weights tables and
adjust the weights accordingly.
The implementation of the reject table, allows us to capture the 
information of prefetches that were rejected, and that can be used 
to further optimize our prefetching mechanism.
%
%\pg{what prior work? this reference is too oblique to be useful,
%  either be specific or drop this paragraph} 
% Prior work in this domain
% lacks the knowledge of which cache blocks should have been prefetched,
% but were not.

@@

\subsection{Optimizing PPF for a Given Prefetcher}
\label{Arch-Generalizing}
The above discussion of PPF shows that it is highly modular and can be
adapted to be used over any {\color{red}underlying} prefetcher for increased prefetch
accuracy. As a first step, all the prefetch
candidates of the prefetcher have to pass through the perceptron filter. 
If qualified, the metadata for perceptron indexing has to be stored. 
Next, when the feedback of a prior prefetch is available in form
of a subsequent demand hit or cache eviction, the stored
metadata needs to be retrieved to update the state of the perceptrons.
% \pg{this isn't quite right, I thought we were taking some metadata
%   from the prefetcher itself like confidence.  You should explicitly
%   mention that prefetcher metadata can be used, update the figure to
%   show that metadata passing from SPP to the prefetcher and point to
%   the figure to show where that connection is}

In general, PPF can be adapted to a {\color{red}new prefetcher} with only a
few modifications:

\noindent \textbf{Making the {\color{red}Underlying} Prefetcher More Aggressive:} By Tuning 
down any internal thresholds or throttling mechanisms to increase its
aggressiveness.

\noindent \textbf{Inferencing and Storing:} All prefetch
recommendations are tested using the perceptron inferencing algorithm.
The perceptron's output, \textit{true} or \textit{false}, should be
saved appropriately, along with all metadata required for perceptron
indexing.

\noindent \textbf{Retrieving and Training:} When feedback for a
prefetch becomes available, the previously stored metadata can be used
to re-index into the perceptron entries and increment or decrement the
weights.

\noindent \textbf{Feature Selection:} Perceptrons essentially
integrate contributions from different features to get a single sum
representing the final confidence.  Thus, perceptron learning can only
be as good as the set of features chosen.  Interestingly, this is what
makes perceptron learning scalable, as it can easily learn to
incorporate newer information in the form of new features.  Some of
the features we developed use information derived directly from
program execution, agnostic to the {\color{red}underlying} prefetcher.  Beyond that,
the feature set can be expanded to convey any useful information or
metadata available in the {\color{red}underlying} prefetcher.

\noindent \textbf{Using Metadata from the Prefetcher:} Some of the 
internal counters specific to the underlying prefetcher can be suitable candidates
for the perceptron features. To make sure that the perceptron layer sees that, 
the relevant metadata must be exported from the prefetcher to PPF. This way,
PPF can be optimized to work tightly-knit with the {\color{red}underlying} prefetcher.

>>>>

Perceptron learning is a light-weight mechanism to pull together
disparate forms of information and synthesize a decision from
them. Our work considers a number of features corresponding to a
prefetch, such as speculation depth, page address and offset, and uses
this information as the inputs to our perceptron-based filter in order
to predict the usefulness of a prefetch.  Here, we discuss our the
design of our proposed perceptron prefetch filter (PPF).  PPF enhances
{an underlying prefetcher} by filtering out predicted unused prefetches. 
PPF is a generalized prefetch filtering mechanism that may be adapted 
to any prefetcher with appropriate feature selection and modifications 
which we describe below.

@@

Figure~\ref{fig:PPF_Datapath} shows the microarchitecture of PPF, as
well as the steps required to filter out not-useful prefetches. The
perceptron filter is organized as a set of tables, where each entry in
the tables holds a \textit{weight}.  For a configuration of PPF using
$N$ number of features, $N$ different tables of weights are needed.
Each feature is used to index a distinct table. The number of entries
of each table varies according to the corresponding feature, hence,
different number of bits are needed to index different tables. Each
weight is a 5-bit saturating counter ranging from -16 to +15. We found
that having 5-bit weights provides a good trade-off between accuracy
and area. A detailed explanation of the storage overhead of PPF can
be found in Section~\ref{Method-Overheads}.
%
%%\pg{the text below would be easier to follow if you used the figure
%%  more. make reference to the figure at the points that you are
%%  talking about}
%
\newline
\newline
\noindent \textbf{Inferencing}\\
%ETG:is this like an explanation of how prefetching happens? 
%EB: I see this as a short context of the datapath leading to the perc filter.
The {prefetcher} is triggered on every demand access to the L2
Cache, as seen in Figure~\ref{fig:PPF_Hierarchy}.  At this point, 
it has the opportunity to trigger a prefetch.
If it does so, it will also need to decide how many cache blocks to
prefetch.  These blocks can be either placed in the L2 or L3 cache
according to the confidence of the prefetching mechanism.  Once the
{underlying} prefetcher is triggered, the suggested prefetch candidates are
fed to the perceptron filter to determine the usefulness of these
prefetches. The filter ultimately decides whether to issue the
prefetch suggestions of the {underlying} prefetcher. As shown in step 1
of Figure~\ref{fig:PPF_Datapath}(a), to make the decision,
each feature corresponding to a suggested prefetch is used to index a
table and all the corresponding weights are summed. The sum denotes
the confidence value for the suggested prefetch, and is thresholded
against two different values: $\tau_{hi}$ and $\tau_{lo}$.

Prefetches whose sum exceeds $\tau_{hi}$ are placed into the L2 cache.
The higher confidence value hints the prefetch would be useful and
should be prioritized.  A prefetch for which the features result in a
confidence value between $\tau_{lo}$ and $\tau_{hi}$ is allocated in
the larger LLC, as the filter is moderately confident of the future
reuse of the cache block, but not enough to possible pollute a
significantly smaller L2.  Suggested prefetches for which the features
lead to a confidence value lower than $\tau_{lo}$ are not prefetched,
as the low confidence value represents that the perceptron learned
that a similar set of features are associated with non-useful
prefetches.
\newline
\newline
\noindent \textbf{Recording}\\
As shown in step 2 of Figure~\ref{fig:PPF_Datapath}(a) the prefetches
that make it through the inference stage are recorded in the
``Prefetch Table''. The prefetch table is a 1,024-entry, direct mapped
structure that contains all metadata required to re-index the
perceptron entries for training. {Ten bits of the address are 
used to index into the tables, and another six bits are stored to 
perform tag matching.}

In addition to the prefetch table mentioned above, PPF also maintains
a 1,024-entry direct-mapped ``Reject Table.'' If a prefetch suggestion
is rejected by the perceptron layer, it is logged into the reject
table. The table is used to train the perceptron to avoid false
negatives \textit{i.e.}, cases where the prediction suggested to
reject the prefetch but the prefetch was proven to be useful based on
the observed demand accesses to the L2.
\newline
\newline
\noindent \textbf{Feedback and Data Retrieval}\\
As depicted in step 3, when there is an eviction or a demand access 
to the L2, training for both the {underlying} prefetcher 
and our filter mechanism is triggered. The address of the cache block 
that triggered the training is used to index both the Prefetch and 
Reject tables. {If it is a match, the corresponding
features are retrieved to index into the tables of perceptron weights.}
%
% Which bits? are the tables accesses sequentially?
%
\newline
\newline
\noindent \textbf{Training}\newline As can be seen in step 4 of, 
the address from the demand
request triggering the training is looked up in both tables. If the
address is in the prefetch table and marked as valid, this hints the
previous prediction was correct and this is a useful prefetch. We
compute the sum of the corresponding weights. If the sum falls below a
specific threshold, training occurs and the corresponding weights are
adjusted accordingly. These thresholds are introduced in order to
avoid over-training, helping the filter adapt quickly to changes in
memory behavior. These thresholds are referred to as $\theta_p$ and
$\theta_n$, respectively for the positive and negative values of
training saturation.

@@

Parallel to accessing the prefetch table, on a demand access, the
reject table is accessed. Before the demand access triggers the next
set of prefetches, the reject table is checked for a valid entry. A
hit means that the corresponding cache block was initially suggested
by the {underlying} prefetcher, but wrongly rejected by the perceptron
filter. The perceptron filter learns from this and makes use of the
corresponding features associated to the original prefetch request,
which are stored in the reject table, to index the weights tables and
adjust the weights accordingly.
The implementation of the reject table, allows us to capture the 
information of prefetches that were rejected, and that can be used 
to further optimize our prefetching mechanism.
%
%\pg{what prior work? this reference is too oblique to be useful,
%  either be specific or drop this paragraph} 
% Prior work in this domain
% lacks the knowledge of which cache blocks should have been prefetched,
% but were not.

@@

\subsection{Optimizing PPF for a Given Prefetcher}
\label{Arch-Generalizing}
The above discussion of PPF shows that it is highly modular and can be
adapted to be used over any {underlying} prefetcher for increased prefetch
accuracy. As a first step, all the prefetch
candidates of the prefetcher have to pass through the perceptron filter. 
If qualified, the metadata for perceptron indexing has to be stored. 
Next, when the feedback of a prior prefetch is available in form
of a subsequent demand hit or cache eviction, the stored
metadata needs to be retrieved to update the state of the perceptrons.
% \pg{this isn't quite right, I thought we were taking some metadata
%   from the prefetcher itself like confidence.  You should explicitly
%   mention that prefetcher metadata can be used, update the figure to
%   show that metadata passing from SPP to the prefetcher and point to
%   the figure to show where that connection is}

\noindent In general, PPF can be adapted to a {new prefetcher} with only a
few modifications:
\newline
\newline
\noindent \textbf{Making the {Underlying} Prefetcher More Aggressive:} By Tuning 
down any internal thresholds or throttling mechanisms to increase its
aggressiveness.
\newline
\newline
\noindent \textbf{Inferencing and Storing:} All prefetch
recommendations are tested using the perceptron inferencing algorithm.
The perceptron's output, \textit{true} or \textit{false}, should be
saved appropriately, along with all metadata required for perceptron
indexing.
\newline
\newline
\noindent \textbf{Retrieving and Training:} When feedback for a
prefetch becomes available, the previously stored metadata can be used
to re-index into the perceptron entries and increment or decrement the
weights.
\newline
\newline
\noindent \textbf{Feature Selection:} Perceptrons essentially
integrate contributions from different features to get a single sum
representing the final confidence.  Thus, perceptron learning can only
be as good as the set of features chosen.  Interestingly, this is what
makes perceptron learning scalable, as it can easily learn to
incorporate newer information in the form of new features.  Some of
the features we developed use information derived directly from
program execution, agnostic to the {underlying} prefetcher.  Beyond that,
the feature set can be expanded to convey any useful information or
metadata available in the {underlying} prefetcher itself.
\newline
\newline
\noindent \textbf{Using Metadata from the Prefetcher:} Some of the 
internal counters specific to the underlying prefetcher can be suitable candidates
for the perceptron features. To make sure that the perceptron layer sees that, 
the relevant metadata must be exported from the prefetcher to PPF. This way,
PPF can be optimized to work tightly-knit with the {underlying} prefetcher.
